---
title: "Step 04: Models - ARIMA and Prophet"
output:
  html_document:
    df_print: paged
---
  
## Setup    
  Load the relevant libraries.
```{r, message=FALSE, warning=FALSE}

# rm(list = ls())
# .rs.restartR()


# data manipulation
library("plyr")
library("tidyverse")
library("magrittr")
library("data.table")
library("lubridate")
library("sqldf")


# time series specific packages
library("timetk")
library("zoo")
library("tibbletime")


# modeling
library("fpp2")
library("prophet")
library("caret")
library("randomForest")
library("xgboost")
library("h2o")
library("keras")
# use_session_with_seed(123456789) # setting the seed to obtain reproducible results
# see https://keras.rstudio.com/articles/faq.html#how-can-i-obtain-reproducible-results-using-keras-during-development and https://cran.r-project.org/web/packages/keras/vignettes/faq.html
# can also re-enable gpu and parallel processing by using:  use_session_with_seed(42, disable_gpu = FALSE, disable_parallel_cpu = FALSE)



# other
library("geosphere")          # specific for distance calculations from lat-lon pairs
library("naniar")             # inspecting missing data
library("rlang")              # building functions
library("recipes")            # used in Keras modeling to design matrices
library("rsample")            # rolling samples for validation stats
library("tfruns")             # used in Keras modeling for trainin runs
library("stringr")            # string manipulation
library("ggplot2")            # viz
library("sweep")              # more easily pull out model statistics
library("yardstick")          # easily calculate accuracy stats
library("doParallel")         # parallel processing

```
  
    
  Session Info.
```{r}

sessionInfo()

```
  
    
  Setup the root directory.
```{r "setup", include = FALSE}

require("knitr")

opts_knit$set(root.dir = "/Users/mdturse/Desktop/Analytics/Chicago_El_Divvy/")

```
  
    
  Setting `wd` as the working directory.
```{r}

wd <- getwd()

wd

```


## Modeling

**NOTE: `period_train`, `period_test`, `skip_span`, `train_data`, `DV_train_data`, and `add_trn_val_test`, `time.Rf.corr_no`, and `time.Xgbtree.corr_yes` are the outputs produced in Step 02 and Step 03**
```{r}

period_train <-
  readRDS(paste0(wd,
                 "/Data/Interim/",
                 "period_train.Rds"
                 )
          )

period_test <-
  readRDS(paste0(wd,
                 "/Data/Interim/",
                 "period_test.Rds"
                 )
          )

skip_span <-
  readRDS(paste0(wd,
                 "/Data/Interim/",
                 "skip_span.Rds"
                 )
          )

train_data <-
  readRDS(paste0(wd,
                 "/Models/",
                 "train_data.Rds"
                 )
          )

DV_train_data <-
  readRDS(paste0(wd,
                 "/Models/",
                 "DV_train_data.Rds"
                 )
          )

add_trn_val_test <-
  readRDS(paste0(wd,
                 "/Data/Interim/",
                 "add_trn_val_test.Rds"
                 )
          )

time.Rf.corr_no <-
  readRDS(paste0(wd,
                 "/Models/",
                 "time.Rf.corr_no.Rds"
                 )
          )

time.Xgbtree.corr_yes <-
  readRDS(paste0(wd,
                 "/Models/",
                 "time.Xgbtree.corr_yes.Rds"
                 )
          )

```  


### Setup for Models Not in `caret` 
  
  This is based on the example shown [here](https://topepo.github.io/rsample/articles/Applications/Time_Series.html), and is needed for the time-series-ish models, which are not currently part of the `caret` modeling process.
    
  Create the rolling-origin resamples to be used for measuring forecast accuracy. Rolling samples will use 2 years of data do predict the next 14 days.  
```{r}

roll_rs <-
  # DV_train_data %>% 
  train_data %>% 
  map(~ rolling_origin(.x,
                       initial = period_train,
                       assess = period_test,
                       cumulative = FALSE,
                       skip = skip_span
                       )
      )


message("DV_train_data")
DV_train_data %>% 
  map(~ nrow(.x)
      )

message("roll_rs")
roll_rs %>% 
  map(~ nrow(.x)
      )


names(roll_rs$`40600`)
roll_rs %>% 
  map(~ length(.x$splits)
      )

# roll_rs$`40600`$splits[[1]]
# train_data %>% 
#   map(~ summarise(.x,
#                   min_d = min(el_date),
#                   max_d = max(el_date),
#                   days = n()
#                   )
#       )

# rm(period_train, period_test, skip_span)

```
  
    
  For plotting, let’s index each split by the first day of the assessment set.
```{r}

get_date <-
  function(x)
    min(assessment(x)$el_date
        )


roll_rs <-
  pmap(.l = list(a = roll_rs),
       .f = function(a) {
         data = a
         
         splits = data$splits %>% 
           map(get_date)
         
         data$start_date = do.call("c", splits)
         
         return(data)
         }
       )

names(roll_rs$`40600`)
length(roll_rs$`40600`$start_date)
head(roll_rs$`40600`$start_date, 20)


rm(get_date)

```



### Model Setup (`forecast::auto.arima` and `prophet::prophet`)
      
  Here, we use `forecast::auto.arima` to produce create an arima model. We also try `prophet::prophet` to create a model based on trend, seasonality, and holidays, and that is a bit like a general additive model. More info can be found [here](https://facebook.github.io/prophet/) and [here](https://peerj.com/preprints/3190/).  
    
  First, I create the function for the basic ARIMA model.
```{r}

fit_model_arima <-
  function(x, ...) {
    data = x %>% 
      analysis() %>% 
      # Since the first day changes over resamples, adjust it based on the first date value in the data frame
      tk_ts(select = el_rides,
            start = .$el_date[[1]] %>% lubridate::year(),
            freq = 7,
            silent = TRUE
            )
    
    fit = auto.arima(data, ...)
    
    return(fit)
    }

```
  
    
  Next I create the function to run an ARIMA model with external regressors (external regressors include fourier transformations, and other regressors identified by using the random forest and xgboost models used above).
```{r}

fit_model_arima_xreg <-
  function(x, ...) {
    data = 
      x %>% 
      analysis() %>% 
      mutate(holiday_binary = if_else(holiday == FALSE,
                                      0,
                                      1
                                      )
             )
    
    # weekly frequency
    ts_7 =
      data %>% 
      tk_ts(select = el_rides,
            start = .$el_date[[1]] %>% lubridate::year(),
            freq = 7,
            silent = TRUE
            )
    
    # yearly frequency
    ts_365 =
      data %>% 
      tk_ts(select = el_rides,
            start = .$el_date[[1]] %>% lubridate::year(),
            # freq = 365.25,
            freq = 365,
            silent = TRUE
            )
    
    # use a fourier transformation to capture daily seasonality and choose K programatically
    bestfit = list(aicc = Inf)
    for(K in seq(7)
        ) {
      n = nrow(assessment(x)
               )
      
      ts_365_fourier = fourier(ts_365,
                               K = K
                               )
      
      ts_365_fourier_future = fourier(ts_365,
                                      K = K,
                                      h = n
                                      )
      
      fit = auto.arima(ts_7,
                       xreg = cbind(ts_365_fourier,
                                    # additional variables identified as "important" with Random Forest and XGBTree models
                                    data$holiday_binary,
                                    data$year,
                                    data$half,
                                    data$quarter,
                                    data$month,
                                    data$mweek,
                                    data$wday.lbl,
                                    data$el_rides_l07,
                                    data$el_rides_l14,
                                    data$el_rides_l21,
                                    data$el_rides_l28,
                                    data$el_rides_ma07,
                                    data$el_rides_ma14,
                                    data$el_rides_ma21,
                                    data$el_rides_ma28
                                    )#,
                       # seasonal = FALSE
                       )
      
      if(fit[["aicc"]] < bestfit[["aicc"]]) {
        bestfit = fit
        bestK = K
        bestts_365_fourier_future = ts_365_fourier_future
      }
      
      return(list(best_fit = bestfit,
                  best_k = bestK,
                  best_ts_365_fourier_future = bestts_365_fourier_future
                  )
             )
    }
    
    return()
    }

```
  
    
  Now I create the function to run the basic prophet model (based just on `el_rides`.
```{r}

# prophet
fit_model_prophet <-
  function(x, ...) {
    x %>% 
      analysis() %>% 
      select(el_date,
             el_rides
             ) %>% 
      rename(ds = el_date,
             y = el_rides
             ) %>% 
      prophet(...)
  }

```


### Run Models

#### Arima-Based Model  
  **Need to review this:**  [http://www.business-science.io/code-tools/2018/04/08/introducing-anomalize.html](http://www.business-science.io/code-tools/2018/04/08/introducing-anomalize.html) for anomaly detection.
    
  Can also use any other modeling methods (e.g., prophet), and then use the outlier detection method twitter used (Generalized ESD) on the remainder. **See:**  [https://www.rdocumentation.org/packages/EnvStats/versions/2.3.0/topics/rosnerTest](https://www.rdocumentation.org/packages/EnvStats/versions/2.3.0/topics/rosnerTest) for the algorithm used outside of `AnomalyDetection::AnomalyDetectionTs`.  
    
  Here I run the basic `forecast::auto.arima` model.
```{r}

# user   system  elapsed 
# 1001.467   21.414 1036.904
# ~ 17 min
message("arima")
start <- proc.time()
models <-
  pmap(.l = list(a = roll_rs),
       .f = function(a) {
         
         splits_a = a$splits %>%
           map(fit_model_arima)
         
         a$arima = splits_a
         
         return(a)
         }
       )

time.arima <- proc.time() - start
time.arima

```
  
    
  Here I run the `forecast::auto.arima` model with external regressors.
```{r}
 
# user  system elapsed 
# 626.467  26.937 673.471
# ~ 11 min
message("arima_xreg")
start <- proc.time()
models <-
  pmap(.l = list(a = models),
       .f = function(a) {
         
         splits_a_xreg = a$splits %>% 
           map(fit_model_arima_xreg)
         
         a$arima_xreg = splits_a_xreg
         
         return(a)
         }
       )

time.arima_xreg <- proc.time() - start
time.arima_xreg

```
  
    
  Here I run the basic `prophet::prophet` model
```{r}

# user  system elapsed 
#  12.007   1.344  13.901
message("prophet")
start <- proc.time()
models <-
  pmap(.l = list(a = models),
       .f = function(a) {
         
         splits_p = a$splits %>%
           map(~ fit_model_prophet(.x,
                                   
                                   )
               )
         
         a$prophet = splits_p
         
         return(a)
         }
       )

time.prophet <- proc.time() - start
time.prophet

```
  
    
  Here I run the `prophet::prophet` model that includes regressors for holidays.
```{r}

# create the holiday_dates
holiday_dates <-
  add_trn_val_test %>% 
  map(~ select(.x,
               holiday_name,
               el_date
               ) %>% 
        filter(holiday_name != "--Not_Holiday--") %>% 
        rename(holiday = holiday_name,
               ds = el_date
               )
      )

holiday_dates <- bind_rows(holiday_dates) %>% 
  select(holiday,
         ds
         ) %>% 
  distinct() %>% 
  arrange(ds)


# user  system elapsed 
#  35.327   3.401  47.577
message("prophet_hol")
start <- proc.time()
models <-
  pmap(.l = list(a = models),
       .f = function(a) {
         
         splits_p_hol = a$splits %>%
           map(~ fit_model_prophet(.x,
                                   holidays = holiday_dates
                                   )
               )
           
         a$prophet_hol <- splits_p_hol
         
         return(a)
         }
       )

time.prophet_hol <- proc.time() - start
time.prophet_hol


# rm(holiday_dates)

```



```{r}

run_times <- list(arima = as.list(time.arima),
                  arima_xreg = as.list(time.arima_xreg),
                  prophet = as.list(time.prophet),
                  prophet_hol = as.list(time.prophet_hol)
                  )

run_times


# saving is done to avoid having to run the models again
saveRDS(models,
        paste0(wd,
               "/Models/",
               "models.Rds"
               )
        )

# models <-
#   readRDS(paste0(wd,
#                  "/Data_Processed/",
#                  "models.Rds"
#                  )
#           )

```
  
    
  Create the `Prophet` forecasts and plots - these will be used for measuring model accuracy (below). Here I creat the future prophet dataset.
```{r}

# Dataframe for future dates
prophet.future <-
  pmap(.l = list(a = models),
       .f = function(a) {
         splits_p_future =
           a$prophet %>% 
           map(make_future_dataframe,
               periods = 365
               )
         
         splits_p_hol_future =
           a$prophet_hol %>% 
           map(make_future_dataframe,
               periods = 365
               )
         
         a$prophet.future <- splits_p_future
         a$prophet_hol.future <- splits_p_hol_future
         
         return(a)
         }
       )

length(prophet.future$`40600`$prophet.future)
length(prophet.future$`40600`$prophet_hol.future)

```  
  
    
  And now I can create the prohpet forecasts. First, the basic prophet forecast.
```{r}

# user  system elapsed 
# 220.704  21.132 242.524
# ~ 4 min
start <- proc.time()
prophet.forecast <-
  pmap(.l = list(a = prophet.future),
       .f = function(a) {
         splits_p_m =
           a$prophet
         
         splits_p_future =
           a$prophet.future
         
         splits_p_forecast =
           pmap(.l = list(b = splits_p_m,
                          c = splits_p_future
                          ),
                .f = function(b, c) {
                  predict(b, c) %>% 
                    # if_else is needed to prevent any negative predictions
                    mutate(yhat_zero_floor = if_else(yhat < 0,
                                                     0,
                                                     yhat
                                                     )
                           )
                  }
                )
         
         a$prophet.forecast = splits_p_forecast
         
         return(a)
         }
       )

time.prophet.forecast <- proc.time() - start
time.prophet.forecast

```
  
    
  And now the prophet forecast including holidays.
```{r}

# user  system elapsed 
# 792.629  78.698 881.271
# ~ 15 min
start <- proc.time()
prophet.forecast <-
  pmap(.l = list(a = prophet.forecast),
       .f = function(a) {
         splits_p_hol_m =
           a$prophet_hol
         
         splits_p_hol_future =
           a$prophet_hol.future
         
         splits_p_hol_forecast =
           pmap(.l = list(b = splits_p_hol_m,
                          c = splits_p_hol_future
                          ),
                .f = function(b, c) {
                  predict(b, c) %>% 
                    # if_else is needed to prevent any negative predictions
                    mutate(yhat_zero_floor = if_else(yhat < 0,
                                                     0,
                                                     yhat
                                                     )
                           )
                  }
                )
         
         a$prophet_hol.forecast = splits_p_hol_forecast
         
         return(a)
         }
       )

time.prophet_hol.forecast <- proc.time() - start
time.prophet_hol.forecast

```
  
    
  Now I simply update the `run_times` dataset with the relevant prophet info, and the relvant info from the random forest and xgboost models.
```{r}


run_times[5:6] <-
  list(prophet.forecast = as.list(time.prophet.forecast),
       prophet_hol.forecast = as.list(time.prophet_hol.forecast)
       )

names(run_times)[5:6] <- c("prophet.forecast", "prophet_hol.forecast")

run_times[7:8] <-
  list(rf_corr_no = as.list(time.Rf.corr_no),
       xgbtree_corr_yes = as.list(time.Xgbtree.corr_yes)
       )
  
names(run_times)[7:8] <- c("rf_corr_no", "xgbtree_corr_yes")


str(run_times)


# saving is done to avoid having to run the forecasts again
saveRDS(run_times,
        paste0(wd,
               "/Models/",
               "run_times.Rds"
               )
        )

# run_times <-
#   readRDS(paste0(wd,
#                  "/Models/",
#                  "run_times.Rds"
#                  )
#           )


rm(time.Rf.corr_no, time.Xgbtree.corr_yes)

```
  
    
  And now I can create the prophet plots.
```{r}

prophet.plots <-
  pmap(.l = list(a = prophet.forecast),
       .f = function(a) {
         splits_p_m =
           a$prophet
         
         splits_p_hol_m =
           a$prophet_hol
         
         splits_p_forecast =
           a$prophet.forecast
         
         splits_p_hol_forecast =
           a$prophet_hol.forecast
         
         
         splits_p_plots =
           pmap(.l = list(b = splits_p_m,
                          c = splits_p_forecast
                          ),
                .f = function(b, c) {
                  plot(b, c)
                  }
                )
         
         splits_p_hol_plots =
           pmap(.l = list(b = splits_p_hol_m,
                          c = splits_p_hol_forecast
                          ),
                .f = function(b, c) {
                  plot(b, c)
                  }
                )
         
         
         a$prophet.plots <- splits_p_plots
         a$prophet_hol.plots <- splits_p_hol_plots
         
         return(a)
         }
       )

# saving is done to avoid having to run the forecasts again
saveRDS(prophet.plots,
        paste0(wd,
               "/Models/",
               "prophet.plots.Rds"
               )
        )

# prophet.plots <-
#   readRDS(paste0(wd,
#                  "/Models/",
#                  "prophet.plots.Rds"
#                  )
#           )


names(prophet.plots)
names(prophet.plots$`40600`)
length(prophet.plots$`40600`$prophet.plots)
length(prophet.plots$`40600`$prophet_hol.plots)

message("prophet")
pmap(.l = list(a = prophet.plots
               ),
     .f = function(a, b) {
       dat = a
       
       dat$prophet.plots[[10]]
       }
     )

message("prophet_hol")
pmap(.l = list(a = prophet.plots
               ),
     .f = function(a, b) {
       dat = a
       
       dat$prophet_hol.plots[[10]]
       
       }
     )

```


