---
title: "Step 06: Models - Accuracy Stats"
output:
  html_document:
    df_print: paged
---
  
## Setup    
  Load the relevant libraries.
```{r, message=FALSE, warning=FALSE}

# rm(list = ls())
# .rs.restartR()


# data manipulation
library("plyr")
library("tidyverse")
library("magrittr")
library("data.table")
library("lubridate")
library("sqldf")


# time series specific packages
library("timetk")
library("zoo")
library("tibbletime")


# modeling
library("fpp2")
library("prophet")
library("caret")
library("randomForest")
library("xgboost")
library("h2o")
library("keras")
# use_session_with_seed(123456789) # setting the seed to obtain reproducible results
# see https://keras.rstudio.com/articles/faq.html#how-can-i-obtain-reproducible-results-using-keras-during-development and https://cran.r-project.org/web/packages/keras/vignettes/faq.html
# can also re-enable gpu and parallel processing by using:  use_session_with_seed(42, disable_gpu = FALSE, disable_parallel_cpu = FALSE)



# other
library("geosphere")          # specific for distance calculations from lat-lon pairs
library("naniar")             # inspecting missing data
library("rlang")              # building functions
library("recipes")            # used in Keras modeling to design matrices
library("rsample")            # rolling samples for validation stats
library("tfruns")             # used in Keras modeling for trainin runs
library("stringr")            # string manipulation
library("ggplot2")            # viz
library("sweep")              # more easily pull out model statistics
library("yardstick")          # easily calculate accuracy stats
library("doParallel")         # parallel processing

```
  
    
  Session Info.
```{r}

sessionInfo()

```
  
    
  Setup the root directory.
```{r "setup", include = FALSE}

require("knitr")

opts_knit$set(root.dir = "/Users/mdturse/Desktop/Analytics/Chicago_El_Divvy/")

```
  
    
  Setting `wd` as the working directory.
```{r}

wd <- getwd()

wd

```


## Modeling

**NOTE: `h2o.limvars_accuracy_stats`, `DV_Fit.Rf.corr_no`, `DV_Fit.Xgbtree.corr_yes`, `func_one_hot_vars`, `run_times`, `h2o.time`, and `h2o.limvars.time` are the outputs produced in Step 02 - Step 05**
```{r}

h2o.limvars_accuracy_stats <-
  readRDS(paste0(wd,
                 "/Models/",
                 "h2o.limvars_accuracy_stats.Rds"
                 )
          )

DV_Fit.Rf.corr_no <-
  readRDS(paste0(wd,
                 "/Models/",
                 "DV_Fit.Rf.corr_no.Rds"
                 )
          )

DV_Fit.Xgbtree.corr_yes <-
  readRDS(paste0(wd,
                 "/Models/",
                 "DV_Fit.Xgbtree.corr_yes.Rds"
                 )
          )

func_one_hot_vars <-
  readRDS(paste0(wd,
                 "/Data/Interim/",
                 "func_one_hot_vars.Rds"
                 )
          )

run_times <-
  readRDS(paste0(wd,
                 "/Models/",
                 "run_times.Rds"
                 )
          )

h2o.time <-
  readRDS(paste0(wd,
                 "/Models/",
                 "h2o.time.Rds"
                 )
          )

h2o.limvars.time <-
  readRDS(paste0(wd,
                 "/Models/",
                 "h2o.limvars.time.Rds"
                 )
          )

```  


## Accuracy Measurements  
    
  Now we can check some summary (RMSE) statistics. **NOTE: these are summary statistics based on the data used to build the models.**  
    
### Interpolation - ARIMA Model
```{r}

accuracy_interp_arima <-
  pmap(.l = list(a = h2o.limvars_accuracy_stats),
       .f = function(a) {
           
         interp_a =
           map_dbl(a$arima,
                   function(x) 
                     sw_glance(x)[["RMSE"]]
                   )
         
         interp_a_xreg =
           map_dbl(a$arima_xreg,
                   function(x) {
                     y = x$best_fit %>% 
                       sw_glance()
                     
                     y[["RMSE"]]
                     }
                   )
         
         a$arima.interpolation = interp_a
         a$arima_xreg.interpolation = interp_a_xreg
         
         return(a)
         }
       )


names(accuracy_interp_arima$`40600`)
accuracy_interp_arima$`40600`$arima.interpolation
accuracy_interp_arima$`40600`$arima_xreg.interpolation

message("arima.interpolation")
accuracy_interp_arima %>% 
  map(~ summary(.x$arima.interpolation)
      )

message("arima_xreg.interpolation")
accuracy_interp_arima %>% 
  map(~ summary(.x$arima_xreg.interpolation)
      )


rm(h2o.limvars_accuracy_stats)

```


### Interpolation - Prophet Model
```{r}

get_interp_prophet <- function(split, pf) {
  # Get analysis data
  pred_dat =
    analysis(split) %>% 
    select(el_date,
           el_rides
         ) %>% 
    left_join(select(pf,
                     ds,
                     yhat,
                     yhat_zero_floor
                     ) %>% 
                mutate_at(vars(ds), as.Date),
              by = c("el_date" = "ds")
            ) %>% 
    mutate(sqrd_error = (el_rides - yhat_zero_floor)^2
           ) %>% 
    arrange(el_date)

  sqrt(mean(pred_dat$sqrd_error,
            na.rm = TRUE
            )
       )
  }    
  

accuracy_interp_prophet <-
  pmap(.l = list(a = accuracy_interp_arima),
       .f = function(a) {
         interp =
           map2_dbl(a$splits,
                    a$prophet.forecast,
                    get_interp_prophet
                    )
         
         interp_hol =
           map2_dbl(a$splits,
                    a$prophet_hol.forecast,
                    get_interp_prophet
                    )
         
         a$prophet.interpolation = interp
         
         a$prophet_hol.interpolation = interp_hol
         
         return(a)
         }
       )


names(accuracy_interp_prophet$`40600`)
accuracy_interp_prophet$`40600`$prophet.interpolation
accuracy_interp_prophet$`40600`$prophet_hol.interpolation

message("prophet.interpolation")
accuracy_interp_prophet %>% 
  map(~ summary(.x$prophet.interpolation)
      )

message("prophet_hol.interpolation")
accuracy_interp_prophet %>% 
  map(~ summary(.x$prophet_hol.interpolation)
      )


# saving is done to avoid having to run the models again
saveRDS(accuracy_interp_prophet,
        paste0(wd,
               "/Models/",
               "accuracy_interp_prophet.Rds"
               )
        )

# accuracy_interp_prophet <-
#   readRDS(paste0(wd,
#                  "/Models/",
#                  "accuracy_interp_prophet.Rds"
#                  )
#           )


rm(accuracy_interp_arima)

```



### Forecast & Interpolation - Random Forest Model  
  
  Get the forecasted data for the random forest and the xgboost tree.  
    
  First, I create a function to get the error percentages.
```{r}

get_pct_error <-
  function(base, mod) {
    data = bind_cols(base,
                     mod
                     ) %>% 
      select(el_date,
             el_rides,
             yhat
             ) %>%
      rename(y = el_rides
             ) %>%
      # if_else is needed to prevent negative predictions
      mutate(yhat_zero_floor = if_else(yhat < 0,
                                       0,
                                       yhat
                                       ),
             sqrd_error = (y - yhat_zero_floor)^2
             ) %>%
      arrange(el_date)
    
    return(data)
    }

```
  
    
  Now I get the forecasts of all ML models as one list.
```{r}

rf_xgb_forecasts_combined <-
  pmap(.l = list(a = accuracy_interp_prophet,
                 b = DV_Fit.Rf.corr_no,
                 c = DV_Fit.Xgbtree.corr_yes
                 ),
       .f = function(a, b, c) {
         splits = a$splits
         rf = b
         xgb = c
         
         fcast =
           pmap(.l = list(d = splits),
                .f = function(d) {
                  # prep the data
                  analysis_assessment =
                    analysis(d) %>% 
                    mutate(type = "interpolation") %>% 
                    bind_rows(assessment(d) %>% 
                                mutate(type = "extrapolation")
                              )
                  
                  one_hot =
                    analysis_assessment %>% 
                    func_one_hot_vars()
                  
                  one_hot_interp =
                    one_hot %>% 
                    filter(typeinterpolation == 1) %>% 
                    select(-typeinterpolation,
                           -typeextrapolation
                           )
                  
                  one_hot_extrap = 
                    one_hot %>% 
                    filter(typeextrapolation == 1) %>% 
                    select(-typeinterpolation,
                           -typeextrapolation
                           )
                  
                  complete_interp =
                    one_hot_interp[complete.cases(one_hot_interp), ]
                  
                  complete_extrap =
                    one_hot_extrap[complete.cases(one_hot_extrap), ]
                  
                  
                  # predictions
                  pred_rf.interp =
                    predict(rf,
                            newdata = complete_interp
                            ) %>% 
                    as.data.frame()
                  
                  names(pred_rf.interp) <- "yhat"
                  
                  pred_rf.extrap =
                    predict(rf,
                            newdata = complete_extrap
                            ) %>% 
                    as.data.frame()
                  
                  names(pred_rf.extrap) <- "yhat"
                  
                  
                  pred_xgb.interp =
                    predict(xgb,
                            newdata = complete_interp
                            ) %>% 
                    as.data.frame()
                  
                  names(pred_xgb.interp) <- "yhat"
                  
                  pred_xgb.extrap =
                    predict(xgb,
                            newdata = complete_extrap
                            ) %>% 
                    as.data.frame()
                  
                  names(pred_xgb.extrap) <- "yhat"
                  
                  
                  # add percent errors
                  full_pred.rf.interp =
                    get_pct_error(base = complete_interp,
                                  mod = pred_rf.interp
                                  )
                  
                  full_pred.rf.extrap =
                    get_pct_error(base = complete_extrap,
                                  mod = pred_rf.extrap
                                  )
                  
                  
                  full_pred.xgb.interp =
                    get_pct_error(base = complete_interp,
                                  mod = pred_xgb.interp
                                  )
                  
                  full_pred.xgb.extrap =
                    get_pct_error(base = complete_extrap,
                                  mod = pred_xgb.extrap
                                  )
                  
                  
                  # result as a list
                  full_pred_list =
                    list(pred.rf.interp = full_pred.rf.interp,
                         pred.rf.extrap = full_pred.rf.extrap,
                         pred.xgb.interp = full_pred.xgb.interp,
                         pred.xgb.extrap = full_pred.xgb.extrap
                         )
                  
                  # return value
                  return(full_pred_list)
                  }
                )
         
         a$ml_forecasts_all = fcast
         
         return(a)
         }
       )

names(rf_xgb_forecasts_combined$`40600`)

```
  
    
  Now I separate the forecats into their own dataframes.
```{r}

rf_xgb_forecasts_individualized <-
  pmap(.l = list(a = rf_xgb_forecasts_combined),
       .f = function(a) {
         fc = a$ml_forecasts_all

         # limit results to the relevant model
         result.rf.interp =
           pmap(.l = list(b = fc),
              .f = function(b) {
                b$pred.rf.interp
                }
              )
         
         result.rf.extrap =
           pmap(.l = list(b = fc),
              .f = function(b) {
                b$pred.rf.extrap
                }
              )
         
         result.xgb.interp =
           pmap(.l = list(b = fc),
              .f = function(b) {
                b$pred.xgb.interp
                }
              )
         
         result.xgb.extrap =
           pmap(.l = list(b = fc),
              .f = function(b) {
                b$pred.xgb.extrap
                }
              )
         
         a$rf.forecast_interp <- result.rf.interp
         a$rf.forecast_extrap <- result.rf.extrap
         a$xgb.forecast_interp <- result.xgb.interp
         a$xgb.forecast_extrap <- result.xgb.extrap
         
         # remove no longer needed list of all forecasts
         a$ml_forecasts_all <- NULL
         
         return(a)
         }
       )


names(rf_xgb_forecasts_individualized$`40600`)


rm(accuracy_interp_prophet, rf_xgb_forecasts_combined)

```
  
    
  Now I calculate the interpolation and extrapolation accuracy stats.
```{r}

# function to get the RMSE value
get_interp_ML <- function(ML.fcast) {
  pred_dat = ML.fcast
  
  sqrt(mean(pred_dat$sqrd_error,
            na.rm = TRUE
            )
       )
  }


accuracy_interp_ML <-
  pmap(.l = list(a = rf_xgb_forecasts_individualized),
       .f = function(a) {
         interp.rf =
           map_dbl(a$rf.forecast_interp,
                   get_interp_ML
                   )
         
         extrap.rf =
           map_dbl(a$rf.forecast_extrap,
                   get_interp_ML
                   )
         
         interp.xgb =
           map_dbl(a$xgb.forecast_interp,
                   get_interp_ML
                   )
         
         extrap.xgb =
           map_dbl(a$xgb.forecast_extrap,
                   get_interp_ML
                   )
         
         a$rf.interpolation <- interp.rf
         a$rf.extrapolation <- extrap.rf
         a$xgb.interpolation <- interp.xgb
         a$xgb.extrapolation <- extrap.xgb
         
         return(a)
         }
       )


names(accuracy_interp_ML$`40600`)
accuracy_interp_ML$`40600`$rf.interpolation
accuracy_interp_ML$`40600`$rf.extrapolation


message("rf.interpolation")
accuracy_interp_ML %>% 
  map(~ summary(.x$rf.interpolation)
      )

message("rf.extrapolation")
accuracy_interp_ML %>% 
  map(~ summary(.x$rf.extrapolation)
      )

message("xgb.interpolation")
accuracy_interp_ML %>% 
  map(~ summary(.x$xgb.interpolation)
      )

message("xgb.extrapolation")
accuracy_interp_ML %>% 
  map(~ summary(.x$xgb.extrapolation)
      )


# saving is done to avoid having to run the models again
saveRDS(accuracy_interp_ML,
        paste0(wd,
               "/Models/",
               "accuracy_interp_ML.Rds"
               )
        )

# accuracy_interp_ML <-
#   readRDS(paste0(wd,
#                  "/Models/",
#                  "accuracy_interp_ML.Rds"
#                  )
#           )


rm(rf_xgb_forecasts_individualized)

```
  
    
  Now we can check some summary (RMSE) statistics. **NOTE: these are summary statistics based on the data NOT used to build the models.**  
    
### Extrapolation - ARIMA Model  
  
  Function to get the ARIMA predictions.
```{r}

get_pred_arima <- function(split, mod) {
  n <- nrow(assessment(split)
            )
  # Get assessment data
  pred_dat <-
    assessment(split) %>%
    mutate(pred = as.vector(forecast(mod, h = n)$mean),
           # if_else is used to put a floor of 0 (zero) for the prediction
           pred_zero_floor = if_else(pred < 0,
                                     0,
                                     pred
                                     ),
           # pct_error = (el_rides - pred_zero_floor) / el_rides * 100,
           sqrd_error = (el_rides - pred_zero_floor)^2
           )
  
  return(pred_dat)
  }

```
  
    
  Function to calculate the extrapolation stats
```{r}

get_extrap_arima <- function(arima.fcast) {
  pred_dat = arima.fcast
  
  sqrt(mean(pred_dat$sqrd_error,
            na.rm = TRUE
            )
       )
  }

```
  
    
  Run the the forecast and extrapolation functions.
```{r}

accuracy_extrap_arima <-
  pmap(.l = list(a = accuracy_interp_ML),
       .f = function(a) {
         pred =
           map2(a$splits,
                a$arima,
                get_pred_arima
               )
         
         a$arima.forecast <- pred
         
         return(a)
         }
       )


accuracy_extrap_arima <-
  pmap(.l = list(a = accuracy_extrap_arima),
       .f = function(a) {
         extrap <-
           map_dbl(a$arima.forecast,
                   get_extrap_arima
                   )
         
         a$arima.extrapolation <- extrap
         
         return(a)
         }
       )

names(accuracy_extrap_arima$`40600`)
accuracy_extrap_arima$`40600`$arima.extrapolation

accuracy_extrap_arima %>% 
  map(~ summary(.x$arima.extrapolation)
      )


rm(accuracy_interp_ML)

```


### Extrapolation - ARIMA Model with XREG  
  
  Function to get the forecasts from the ARIMA XREG models. 
```{r}

get_pred_arima_xreg <-
  function(split, mod) {
    n = nrow(assessment(split)
             )
    
    data_og =
      split %>% 
      assessment %>% 
      mutate(holiday_binary = if_else(holiday == FALSE,
                                      0,
                                      1
                                      )
             )
    
    data_pred =
      data_og %>% 
      mutate(pred = as.vector(forecast(mod$best_fit,
                                       xreg = cbind(mod$best_ts_365_fourier_future,
                                                    # variables below identified based on "importance" from Random Forest and XGBTree models
                                                    data_og$holiday_binary,
                                                    data_og$year,
                                                    data_og$half,
                                                    data_og$quarter,
                                                    data_og$month,
                                                    data_og$mweek,
                                                    data_og$wday.lbl,
                                                    data_og$el_rides_l07,
                                                    data_og$el_rides_l14,
                                                    data_og$el_rides_l21,
                                                    data_og$el_rides_l28,
                                                    data_og$el_rides_ma07,
                                                    data_og$el_rides_ma14,
                                                    data_og$el_rides_ma21,
                                                    data_og$el_rides_ma28
                                    ),
                                       h = n
                                       )$mean
                              ),
             # if_else is used to put a floor of 0 (zero) for the prediction
             pred_zero_floor = if_else(pred < 0,
                                       0,
                                       pred
                                       ),
             # pct_error = (el_rides - pred_zero_floor) / el_rides * 100
             sqrd_error = (el_rides - pred_zero_floor)^2
             )
    }

```
  
    
  Function to calculate the extrapolation stats.
```{r}

get_extrap_arima_xreg <-
  function(arima.fcast) {
    pred_dat = arima.fcast

    sqrt(mean(pred_dat$sqrd_error,
              na.rm = TRUE
              )
         )
    }

```
  
    
  Run the forecast and extrapolation functions.
```{r}

accuracy_extrap_arima_xreg <-
  pmap(.l = list(a = accuracy_extrap_arima),
       .f = function(a) {
         pred <-
           map2(a$splits,
                a$arima_xreg,
                get_pred_arima_xreg
               )
         
         a$arima_xreg.forecast <- pred
         
         return(a)
         }
       )


accuracy_extrap_arima_xreg <-
  pmap(.l = list(a = accuracy_extrap_arima_xreg),
       .f = function(a) {
         extrap <-
           map_dbl(a$arima_xreg.forecast,
                   get_extrap_arima
                   )
         
         a$arima_xreg.extrapolation <- extrap
         
         return(a)
         }
       )

names(accuracy_extrap_arima_xreg$`40600`)
accuracy_extrap_arima_xreg$`40600`$arima_xreg.extrapolation

accuracy_extrap_arima_xreg %>% 
  map(~ summary(.x$arima_xreg.extrapolation)
      )


rm(accuracy_extrap_arima)


```


### Extrapolation - Prophet Model  
  
  Function to get the extrapolation stats.
```{r}

get_extrap_prophet <- function(split, pf) {
  # Get assessment data
  pred_dat <-
    assessment(split) %>% 
    select(el_date,
           el_rides
         ) %>% 
    left_join(select(pf,
                     ds,
                     yhat,
                     yhat_zero_floor
                     ) %>% 
                mutate_at(vars(ds), as.Date),
              by = c("el_date" = "ds")
            ) %>% 
    mutate(sqrd_error = (el_rides - yhat_zero_floor)^2 
           ) %>% 
    arrange(el_date)

  sqrt(mean(pred_dat$sqrd_error,
            na.rm = TRUE
            )
       )
  }    
  
```
  
    
  Run the extrapolation stats function.
```{r}

accuracy_extrap_prophet <-
  pmap(.l = list(a = accuracy_extrap_arima_xreg),
       .f = function(a) {
         extrap <-
           map2_dbl(a$splits,
                    a$prophet.forecast,
                    get_extrap_prophet
                    )
         
         extrap_hol <-
           map2_dbl(a$splits,
                    a$prophet_hol.forecast,
                    get_extrap_prophet
                    )
         
         a$prophet.extrapolation <- extrap
         
         a$prophet_hol.extrapolation <- extrap_hol
         
         return(a)
         }
       )


# saving is done to avoid having to run the models again
saveRDS(accuracy_extrap_prophet,
        paste0(wd,
               "/Models/",
               "accuracy_extrap_prophet.Rds"
               )
        )

# accuracy_extrap_prophet <-
#   readRDS(paste0(wd,
#                  "/Models/",
#                  "accuracy_extrap_prophet.Rds"
#                  )
#           )

names(accuracy_extrap_prophet$`40600`)
accuracy_extrap_prophet$`40600`$prophet.extrapolation
accuracy_extrap_prophet$`40600`$prophet_hol.extrapolation

message("prophet")
accuracy_extrap_prophet %>% 
  map(~ summary(.x$prophet.extrapolation)
      )

message("prophet_hol")
accuracy_extrap_prophet %>% 
  map(~ summary(.x$prophet_hol.extrapolation)
      )


rm(accuracy_extrap_arima_xreg)

```
  
    
  Add an identifier for the relevant `el_stop_id` value.
```{r}
  
add_el_stop_id <-
  pmap(.l = list(a = accuracy_extrap_prophet,
                 c = names(accuracy_extrap_prophet)
                 ),
       .f = function(a, c) {
         a$el_stop_id <- c
         
         return(a)
       }
       )

# saving is done to avoid having to run the forecasts again
saveRDS(add_el_stop_id,
        paste0(wd,
               "/Models/",
               "add_el_stop_id.Rds"
               )
        )

# add_el_stop_id <-
#   readRDS(paste0(wd,
#                  "/Models/",
#                  "add_el_stop_id.Rds"
#                  )
#           )


names(add_el_stop_id$`40600`)
names(add_el_stop_id$`40910`)


rm(accuracy_extrap_prophet)

```
  
    
  Now I simply update the `run_times` dataset with the relevant h2o info.
```{r}


run_times[9:10] <-
  list(h2o = as.list(h2o.time),
       h2o.limvars = as.list(h2o.limvars.time)
       )

names(run_times)[9:10] <- c("h2o", "h2o.limvars")


str(run_times)


# saving is done to avoid having to run the forecasts again
saveRDS(run_times,
        paste0(wd,
               "/Models/",
               "run_times.Rds"
               )
        )

# run_times <-
#   readRDS(paste0(wd,
#                  "/Models/",
#                  "run_times.Rds"
#                  )
#           )

```


