---
title: "Step 05: Models - H20"
output:
  html_document:
    df_print: paged
---
  
## Setup    
  Load the relevant libraries.
```{r, message=FALSE, warning=FALSE}

# rm(list = ls())
# .rs.restartR()


# data manipulation
library("plyr")
library("tidyverse")
library("magrittr")
library("data.table")
library("lubridate")
library("sqldf")


# time series specific packages
library("timetk")
library("zoo")
library("tibbletime")


# modeling
library("fpp2")
library("prophet")
library("caret")
library("randomForest")
library("xgboost")
library("h2o")
library("keras")
# use_session_with_seed(123456789) # setting the seed to obtain reproducible results
# see https://keras.rstudio.com/articles/faq.html#how-can-i-obtain-reproducible-results-using-keras-during-development and https://cran.r-project.org/web/packages/keras/vignettes/faq.html
# can also re-enable gpu and parallel processing by using:  use_session_with_seed(42, disable_gpu = FALSE, disable_parallel_cpu = FALSE)



# other
library("geosphere")          # specific for distance calculations from lat-lon pairs
library("naniar")             # inspecting missing data
library("rlang")              # building functions
library("recipes")            # used in Keras modeling to design matrices
library("rsample")            # rolling samples for validation stats
library("tfruns")             # used in Keras modeling for trainin runs
library("stringr")            # string manipulation
library("ggplot2")            # viz
library("sweep")              # more easily pull out model statistics
library("yardstick")          # easily calculate accuracy stats
library("doParallel")         # parallel processing

```
  
    
  Session Info.
```{r}

sessionInfo()

```
  
    
  Setup the root directory.
```{r "setup", include = FALSE}

require("knitr")

opts_knit$set(root.dir = "/Users/mdturse/Desktop/Analytics/Chicago_El_Divvy/")

```
  
    
  Setting `wd` as the working directory.
```{r}

wd <- getwd()

wd

```


## Modeling

**NOTE: `add_trn_val_test`, `func_one_hot_vars`, `prophet.plots`, and `DV_corr_predict` are the outputs produced in Step 02 - Step 04**
```{r}

add_trn_val_test <-
  readRDS(paste0(wd,
                 "/Data/Interim/",
                 "add_trn_val_test.Rds"
                 )
          )

func_one_hot_vars <-
  readRDS(paste0(wd,
                 "/Data/Interim/",
                 "func_one_hot_vars.Rds"
                 )
          )

prophet.plots <-
  readRDS(paste0(wd,
                 "/Models/",
                 "prophet.plots.Rds"
                 )
          )

DV_corr_predict <-
  readRDS(paste0(wd,
                 "/Data/Interim/",
                 "DV_corr_predict.Rds"
                 )
          )

```  


## Modeling With H2O


### All Variables  
  
  Fire up h2o, but turn off the progress bar.
```{r}

h2o.init()
h2o.no_progress()

```
  
    
  Now, create one-hot variables from the `add_trn_val_test` dataset, then separate the data into train, validation, and test datasets.
```{r}

# One-Hot
h2o_one_hot <-
  add_trn_val_test %>% 
  map(~ func_one_hot_vars(.x))


# create train-val-test for h2o
h2o_train_tbl <-
  h2o_one_hot %>% 
  map(~ (filter(.x,
                data_use_el_stop_id == "01_train"
                )
         )
      )

h2o_valid_tbl <-
  h2o_one_hot %>% 
  map(~ (filter(.x,
                data_use_el_stop_id == "02_validation"
                )
         )
      )

h2o_test_tbl <-
  h2o_one_hot %>% 
  map(~ (filter(.x,
                data_use_el_stop_id == "03_test"
                )
         )
      )


# Convert to H2OFrame objects
h2o_train <- h2o_train_tbl %>% map(~ as.h2o(.x))
h2o_valid <- h2o_valid_tbl %>% map(~ as.h2o(.x))
h2o_test  <- h2o_test_tbl %>% map(~ as.h2o(.x))

```
  
    
  Create the Interpolation & Extrapolation Datasets
```{r}

h2o_data <-
  pmap(.l = list(a = prophet.plots),
       .f = function(a) {
         dat_interp = a$splits %>% 
           map(~ analysis(.x) %>% 
                 func_one_hot_vars() %>% 
                 as.h2o()
               )
         
         dat_extrap = a$splits %>% 
           map(~ assessment(.x) %>% 
                 func_one_hot_vars() %>% 
                 as.h2o()
               )
         
         a$interp_data = dat_interp
         a$extrap_data = dat_extrap
         
         return(a)
         }
       )

# h2o_data$`40600`$interp_data[[1]] %>% as.data.frame() %>% dim()
rm(prophet.plots)

```
  
    
  Set the names for h2o, and run the `h2o.automl`.
```{r}

# Set names for h2o
y <- "el_rides"
x <- h2o_train %>% 
  map(~ setdiff(names(.x),
                y
                )
      )


# user  system elapsed 
#   9.730   5.819 215.764
# ~ 4 min
start <- proc.time()

h2o_automl_models <-
  pmap(.l = list(a = x,
                 b = h2o_train,
                 c = h2o_valid,
                 d = h2o_test
                 ),
       .f = function(a, b, c, d) {
         h2o.automl(
           x = a,
           y = y,
           training_frame = b,
           validation_frame = c,
           leaderboard_frame = d,
           nfolds = 13,
           max_runtime_secs = 30,
           stopping_metric = "deviance"
           )
         }
       )

h2o.time <- proc.time() - start
rm(start)

h2o.time


saveRDS(h2o.time,
        paste0(wd,
               "/Models/",
               "h2o.time.Rds"
               )
        )

# h2o.time <-
#   readRDS(paste0(wd,
#                  "/Models/",
#                  "h2o.time.Rds"
#                  )
#           )

# print(automl_models_h2o$`40600`@leaderboard)

```
  
    
  Extract leader model.
```{r}

h2o_automl_leader <-
  pmap(.l = list(a = h2o_automl_models),
       .f = function(a) {
         ldr = a@leader
         
         return(ldr)
         }
       )


# save the leader
saveRDS(h2o_automl_leader,
        paste0(wd,
               "/Models/",
               "h2o_automl_leader.Rds"
               )
        )

# h2o_automl_leader <-
#   readRDS(paste0(wd,
#                  "/Models/",
#                  "h2o_automl_leader.Rds"
#                  )
#           )

```
  
    
  Get the predictions.
```{r}

# predictions on test set
h2o_pred <-
  pmap(.l = list(a = h2o_automl_leader,
                 b = h2o_test
                 ),
       .f = function(a, b) {
         res = h2o.predict(a, newdata = b)
         
         return(res)
         }
       )


# predictions on interp & extrap
h2o_forecasts <-
  pmap(.l = list(a = h2o_data,
                 b = h2o_automl_leader
                 ),
       .f = function(a, b) {
         int_pred =
           pmap(.l = list(c = a$interp_data),
                .f = function(c) {
                  pred = h2o.predict(b, newdata = c) %>% 
                    as.data.frame()
                  
                  return(pred)
                  }
                )

         ext_pred =
           pmap(.l = list(d = a$extrap_data),
                .f = function(d) {
                  pred = h2o.predict(b, newdata = d) %>% 
                    as.data.frame()
                  
                  return(pred)
                  }
                )
         
         
         int_data =
           pmap(.l = list(e = a$interp_data),
                .f = function(e) {
                  df = e %>% as.data.frame()
                  
                  return(df)
                  }
                )
         
         ext_data =
           pmap(.l = list(f = a$extrap_data),
                .f = function(f) {
                  df = f %>% as.data.frame()
                  
                  return(df)
                  }
                )
         
         
         int_fcast =
           pmap(.l = list(g = int_data,
                          h = int_pred
                          ),
                .f = function(g, h) {
                  res = g %>% 
                    select(el_date, el_rides) %>% 
                    bind_cols(h) %>% 
                    rename(actual = el_rides,
                           yhat = predict
                           ) %>% 
                    mutate(sqrd_error = (actual - yhat)^2
                           )
                  }
                )
         
         ext_fcast =
           pmap(.l = list(i = ext_data,
                          j = ext_pred
                          ),
                .f = function(i, j) {
                  res = i %>% 
                    select(el_date, el_rides) %>% 
                    bind_cols(j) %>% 
                    rename(actual = el_rides,
                           yhat = predict
                           ) %>% 
                    mutate(sqrd_error = (actual - yhat)^2
                           )
                  }
                )
         
         
         a$h2o.interp.forecast = int_fcast
         a$h2o.extrap.forecast = ext_fcast
         
         return(a)
         }
       )


# add_el_stop_id$`40600`$h2o.extrap.forecast[[1]] %>% str()
rm(h2o_data)

```
  
    
  Now I can calculate the accuracy (RMSE) stats.
```{r}

# calculation of RMSE
h2o_accuracy_stats <-
  pmap(.l = list(a = h2o_forecasts),
       .f = function(a) {
         rmse_interp =
           a$h2o.interp.forecast %>% 
           map_dbl(~ sqrt(mean(.x$sqrd_error)
                          )
               )
         
         rmse_extrap =
           a$h2o.extrap.forecast %>% 
           map_dbl(~ sqrt(mean(.x$sqrd_error)
                          )
               )
         
         a$h2o.interpolation = rmse_interp
         a$h2o.extrapolation = rmse_extrap
         
         return(a)
         }
       )


# save the dataset 
saveRDS(h2o_accuracy_stats,
        paste0(wd,
               "/Models/",
               "h2o_accuracy_stats.Rds"
               )
        )


# h2o_accuracy_stats <-
#   readRDS(paste0(wd,
#                  "/Models/",
#                  "h2o_accuracy_stats.Rds"
#                  )
#           )


rm(h2o_forecasts)


# Summary stats of accuracy  
h2o_accuracy_stats %>% 
  map(~ summary(.x$h2o.interpolation)
      )

h2o_accuracy_stats %>% 
  map(~ summary(.x$h2o.extrapolation)
      )

```
  
    
  Now I measure performance on the test dataset.
```{r}

h2o_perf <-
  pmap(.l = list(a = h2o_automl_leader,
                 b = h2o_test),
       .f = function(a, b) {
         per = h2o.performance(a, newdata = b)
         
         return(per)
         }
       )

h2o_perf


# save the leader performance
saveRDS(h2o_perf,
        paste0(wd,
               "/Models/",
               "h2o_perf.Rds"
               )
        )


# h2o_perf <-
#   readRDS(paste0(wd,
#                  "/Models/",
#                  "h2o_perf.Rds"
#                  )
#           )


# h2o.removeAll()

```


### Limited Variables  
  
  Select specific columns to use
```{r}

col_names <-
  DV_corr_predict %>% 
  map(~ colnames(.x))


select_cols <-
  pmap(.l = list(a = h2o_one_hot,
                 b = col_names
                 ),
       .f = function(a, b) {
         res = a[, colnames(a) %in% b]
         
         return(res)
         }
       )


rm(col_names)

```
  
    
  Create train-val-test datasets for h2o.
```{r}

h2o.limvars_train_tbl <-
  select_cols %>% 
  map(~ (filter(.x,
                data_use_el_stop_id == "01_train"
                )
         )
      )

h2o.limvars_valid_tbl <-
  select_cols %>% 
  map(~ (filter(.x,
                data_use_el_stop_id == "02_validation"
                )
         )
      )

h2o.limvars_test_tbl <-
  select_cols %>% 
  map(~ (filter(.x,
                data_use_el_stop_id == "03_test"
                )
         )
      )


rm(select_cols)


# Convert to H2OFrame objects
h2o.limvars_train <- h2o.limvars_train_tbl %>% map(~ as.h2o(.x))
h2o.limvars_valid <- h2o.limvars_valid_tbl %>% map(~ as.h2o(.x))
h2o.limvars_test  <- h2o.limvars_test_tbl %>% map(~ as.h2o(.x))

```
  
    
  Set names for h2o and run `h2o.automl`.
```{r}

y <- "el_rides"
x <- h2o.limvars_train %>% 
  map(~ setdiff(names(.x),
                y
                )
      )


# user  system elapsed 
#   7.985   3.463 208.442 
# ~ 4 min
start <- proc.time()

h2o.limvars_automl_models <-
  pmap(.l = list(a = x,
                 b = h2o.limvars_train,
                 c = h2o.limvars_valid,
                 d = h2o.limvars_test
                 ),
       .f = function(a, b, c, d) {
         h2o.automl(
           x = a,
           y = y,
           training_frame = b,
           validation_frame = c,
           leaderboard_frame = d,
           nfolds = 13,
           # max_runtime_secs = 300,
           max_runtime_secs = 30,
           stopping_metric = "deviance"
           )
         }
       )

h2o.limvars.time <- proc.time() - start
rm(start)

h2o.limvars.time


saveRDS(h2o.limvars.time,
        paste0(wd,
               "/Models/",
               "h2o.limvars.time.Rds"
               )
        )

# h2o.limvars.time <-
#   readRDS(paste0(wd,
#                  "/Models/",
#                  "h2o.limvars.time.Rds"
#                  )
#           )

# print(h2o.limvars_automl_models$`40600`@leaderboard)

```
  
    
  Extract leader model.
```{r}

h2o.limvars_automl_leader <-
  pmap(.l = list(a = h2o.limvars_automl_models),
       .f = function(a) {
         ldr = a@leader
         
         return(ldr)
         }
       )


# save the leader
saveRDS(h2o.limvars_automl_leader,
        paste0(wd,
               "/Models/",
               "h2o.limvars_automl_leader.Rds"
               )
        )

# h2o.limvars_automl_leader <-
#   readRDS(paste0(wd,
#                  "/Models/",
#                  "h2o.limvars_automl_leader.Rds"
#                  )
#           )

```
  
    
  Get the predictions on the interpolation, and extrapolation, and test data.
```{r}

# predictions on interp & extrap
h2o.limvars_forecasts <-
  pmap(.l = list(a = h2o_accuracy_stats,
                 b = h2o.limvars_automl_leader
                 ),
       .f = function(a, b) {
         int_pred =
           pmap(.l = list(c = a$interp_data),
                .f = function(c) {
                  pred = h2o.predict(b, newdata = c) %>% 
                    as.data.frame()
                  
                  return(pred)
                  }
                )
         
         ext_pred =
           pmap(.l = list(d = a$extrap_data),
                .f = function(d) {
                  pred = h2o.predict(b, newdata = d) %>% 
                    as.data.frame()
                  
                  return(pred)
                  }
                )
         
         
         int_data =
           pmap(.l = list(e = a$interp_data),
                .f = function(e) {
                  df = e %>% as.data.frame()
                  
                  return(df)
                  }
                )
         
         ext_data =
           pmap(.l = list(f = a$extrap_data),
                .f = function(f) {
                  df = f %>% as.data.frame()
                  
                  return(df)
                  }
                )
         
         
         int_fcast =
           pmap(.l = list(g = int_data,
                          h = int_pred
                          ),
                .f = function(g, h) {
                  res = g %>% 
                    select(el_date, el_rides) %>% 
                    bind_cols(h) %>% 
                    rename(actual = el_rides,
                           yhat = predict
                           ) %>% 
                    mutate(sqrd_error = (actual - yhat)^2
                           )
                  }
                )
         
         ext_fcast =
           pmap(.l = list(i = ext_data,
                          j = ext_pred
                          ),
                .f = function(i, j) {
                  res = i %>% 
                    select(el_date, el_rides) %>% 
                    bind_cols(j) %>% 
                    rename(actual = el_rides,
                           yhat = predict
                           ) %>% 
                    mutate(sqrd_error = (actual - yhat)^2
                           )
                  }
                )
         
         
         a$h2o.limvars.interp.forecast = int_fcast
         a$h2o.limvars.extrap.forecast = ext_fcast
         
         return(a)
         }
       )


rm(h2o_accuracy_stats)


# predictions on test set
h2o.limvars_pred <-
  pmap(.l = list(a = h2o.limvars_automl_leader,
                 b = h2o.limvars_test
                 ),
       .f = function(a, b) {
         res = h2o.predict(a, newdata = b)
         
         return(res)
         }
       )

```
  
    
  Calculate the accuracy (RMSE) stats.
```{r}

# calculate RMSE
h2o.limvars_accuracy_stats <-
  pmap(.l = list(a = h2o.limvars_forecasts),
       .f = function(a) {
         rmse_interp =
           a$h2o.limvars.interp.forecast %>% 
           map_dbl(~ sqrt(mean(.x$sqrd_error)
                          )
               )
         
         rmse_extrap =
           a$h2o.limvars.extrap.forecast %>% 
           map_dbl(~ sqrt(mean(.x$sqrd_error)
                          )
               )
         
         a$h2o.limvars.interpolation = rmse_interp
         a$h2o.limvars.extrapolation = rmse_extrap
         
         return(a)
         }
       )


# save the dataset 
saveRDS(h2o.limvars_accuracy_stats,
        paste0(wd,
               "/Models/",
               "h2o.limvars_accuracy_stats.Rds"
               )
        )


# h2o.limvars_accuracy_stats <-
#   readRDS(paste0(wd,
#                  "/Models/",
#                  "h2o.limvars_accuracy_stats.Rds"
#                  )
#           )


rm(h2o.limvars_forecasts)


# Summary stats of accuracy  
h2o.limvars_accuracy_stats %>% 
  map(~ summary(.x$h2o.limvars.interpolation)
      )

h2o.limvars_accuracy_stats %>% 
  map(~ summary(.x$h2o.limvars.extrapolation)
      )

```
  
    
  Measure performance on test dataset.
```{r}

h2o.limvars_perf <-
  pmap(.l = list(a = h2o.limvars_automl_leader,
                 b = h2o.limvars_test),
       .f = function(a, b) {
         per = h2o.performance(a, newdata = b)
         
         return(per)
         }
       )

h2o.limvars_perf


# save the leader performance
saveRDS(h2o.limvars_perf,
        paste0(wd,
               "/Models/",
               "h2o.limvars_perf.Rds"
               )
        )


# h2o.limvars_perf <-
#   readRDS(paste0(wd,
#                  "/Models/",
#                  "h2o.limvars_perf.Rds"
#                  )
#           )


# h2o.removeAll()

```


